---
title: "QBS122 Project"
output:
  html_document: default
  pdf_document: default
date: "2025-06-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(igraph)
library(tidyverse)
library(leaflet)
library(geosphere)
```
Read the node and edges of the cvs file, using the function in tidyverse
```{r}
nodes_df <- read.csv("/Users/henri/Desktop/QBS122 Project/Data/Reduced_node.csv", stringsAsFactors = FALSE, check.names = FALSE)
edges_df <- read.csv("/Users/henri/Desktop/QBS122 Project/Data/Reduced_edge.csv", stringsAsFactors = FALSE, check.names = FALSE)
head(nodes_df)
head(edges_df)
```

Create the social network graph we need 
```{r}
g <- graph_from_data_frame(
  d = edges_df %>% rename(from = Source, to = Target),
  vertices = nodes_df %>% rename(name = Id),
  directed = FALSE
)

V(g)$label   <- nodes_df$Label       # Add a "label" attribute to each vertex
V(g)$owner   <- nodes_df$Owner
V(g)$role    <- nodes_df$Role
V(g)$lon     <- nodes_df$Longitude
V(g)$lat     <- nodes_df$Latitude
```

For the data I want to give you a initial plot of the data, and then we will be able to do the 
```{r}
set.seed(42) 
layout_mat <- layout_with_fr(g)
plot(
  g,
  layout = layout_mat,
  vertex.size = 5,
  vertex.label = NA, 
  vertex.color = "skyblue",
  edge.color   = "gray80"
)

```
Let us see where is the real picture, we use the pakage geosphere to draw the actual graph
```{r}
coords <- data.frame(
  id    = V(g)$name,
  lon   = as.numeric(V(g)$lon),
  lat   = as.numeric(V(g)$lat),
  label = V(g)$label,
  stringsAsFactors = FALSE
)

min_lon <- min(coords$lon, na.rm = TRUE)
max_lon <- max(coords$lon, na.rm = TRUE)
min_lat <- min(coords$lat, na.rm = TRUE)
max_lat <- max(coords$lat, na.rm = TRUE)

m <- leaflet(data = coords) %>%
  addTiles()

m <- m %>%
  addCircleMarkers(
    lng        = ~lon,
    lat        = ~lat,
    label      = ~label,  
    radius     = 1,
    color      = "dodgerblue",
    fillOpacity= 0.7
  )
m
```
First Itry to plot it with the default ways
```{r}
layout_geo <- cbind(
  as.numeric(V(g)$lon),
  as.numeric(V(g)$lat)
)

plot(
  g,
  layout       = layout_geo,
  asp          = 1, 
  vertex.size  = 5,       
  vertex.label = NA,        
  vertex.color = "skyblue",
  edge.color   = "gray80",
  main         = "Network Plotted on Geographic Coordinates"
)
```

For the question 3, we compute the Global Network Metrics
```{r}
num_nodes <- vcount(g)
num_edges <- ecount(g)
dens       <- edge_density(g, loops = FALSE)
avg_path   <- mean_distance(g, directed = FALSE, unconnected = TRUE)
net_diam   <- diameter(g, directed = FALSE, unconnected = TRUE)
clust_glob <- transitivity(g, type = "global")
clust_loc  <- transitivity(g, type = "local")
avg_clust  <- mean(clust_loc, na.rm = TRUE)
avg_deg    <- mean(degree(g))
deg_vals   <- degree(g)
deg_summary<- summary(deg_vals)

comm             <- cluster_louvain(g)
V(g)$community   <- membership(comm)
comm_sizes       <- sizes(comm)

# Print out a summary of network-level metrics
cat("—— Network Summary Metrics ——\n")
cat("Number of nodes: ", num_nodes, "\n")
cat("Number of edges: ", num_edges, "\n")
cat("Density: ", round(dens, 4), "\n")
cat("Average path length: ", round(avg_path, 4), "\n")
cat("Diameter: ", net_diam, "\n")
cat("Global clustering coefficient: ", round(clust_glob, 4), "\n")
cat("Average local clustering coefficient: ", round(avg_clust, 4), "\n")
cat("Average degree: ", round(avg_deg, 4), "\n")
cat("Degree distribution summary:\n"); print(deg_summary)
cat("Number of communities detected: ", length(comm_sizes), "\n")
cat("Sizes of each community:\n"); print(comm_sizes)

```

```{r}
deg_cent       <- degree(g, normalized = TRUE)

# Betweenness centrality (normalized to [0,1])
betw_cent      <- betweenness(g, normalized = TRUE)

# Closeness centrality (normalized to [0,1])
closeness_cent <- closeness(g, normalized = TRUE)

# Store these centralities back into the graph’s vertex attributes
V(g)$degree_c      <- deg_cent
V(g)$betweenness_c <- betw_cent
V(g)$closeness_c   <- closeness_cent

# Build a data frame of node‐level statistics and attributes
node_positions_df <- data.frame(
  id            = V(g)$name,
  label         = V(g)$label,
  degree_c      = deg_cent,
  betweenness_c = betw_cent,
  closeness_c   = closeness_cent,
  community     = V(g)$community,
  stringsAsFactors = FALSE
)

# Show the top 5 nodes by degree centrality
cat("—— Node‐Level Centrality Rankings (Top 5) ——\n")
cat("Top 5 by Degree Centrality:\n")
print(
  node_positions_df %>%
    arrange(desc(degree_c)) %>%
    slice(1:5)
)
cat("Top 5 by Betweenness Centrality:\n")
print(
  node_positions_df %>%
    arrange(desc(betweenness_c)) %>%
    slice(1:5)
)
```
```{r}
# —— D. Compute a layout and plot the network —— 

set.seed(2025)  # For reproducibility of the layout
layout_mat <- layout_with_fr(g)  # Fruchterman–Reingold force‐directed layout

# Store the x–y coordinates as vertex attributes
V(g)$x_coord <- layout_mat[,1]
V(g)$y_coord <- layout_mat[,2]

# Add these coordinates to our node_positions_df
node_positions_df$x <- V(g)$x_coord
node_positions_df$y <- V(g)$y_coord

# Basic scatter‐plot visualization
plot(
  node_positions_df$x,
  node_positions_df$y,
  pch  = 21,
  bg   = ifelse(node_positions_df$community == 1, "tomato", "skyblue"),  
        # use color “tomato” for community 1, “skyblue” for all others (as an example)
  cex  = node_positions_df$degree_c * 10 + 0.5,  
        # scale each point’s size by its (normalized) degree‐centrality
  xlab = "X (Fruchterman–Reingold)",
  ylab = "Y (Fruchterman–Reingold)",
  main = "Network Layout (FR)\nNode Size ∝ Degree Centrality, Color ∝ Community"
)

# Label the top 5 highest‐degree nodes
top5_deg_nodes <- node_positions_df %>%
  arrange(desc(degree_c)) %>%
  slice(1:5)

text(
  top5_deg_nodes$x,
  top5_deg_nodes$y,
  labels = top5_deg_nodes$label,
  pos    = 3,       # position text above the point
  cex    = 0.8,
  col    = "darkred"
)

```






